% $Id: install.tex,v 1.1 2002/02/19 06:45:55 mkdist Exp $
\chapter{Installation Guide}

This chapter describes how to get BProc built and installed on your
system.  This process is fairly involved.  If you are new to BProc, it
is recommended that you try a pre-built set of kernel, BProc and LANL
beoboot packages from
\texttt{http://www.clustermatic.org}.

\subsection*{Build a Linux kernel with BProc support}
First you will have to build a Linux kernel with the BProc kernel
patch applied.  The kernel patch is provied with the BProc source
distribution in the \texttt{patches} directory.  The patch file is
named \texttt{bproc-patch-\emph{x.y.z}} where \emph{x.y.z} is the
kernel revision the patch is built against.  The patch may apply
cleanly against other revisions which are close to the one indicated.
The kernel patches are built against stock Linux kernels.  They will
most likely not apply cleanly to the kernels distributed by vendors
such as Red Hat even if their kernel has the same revision number.

The following example shows how you might apply the patch to your
linux kernel tree:
\begin{verbatim}
cd /usr/src/linux
patch -p1 < ~/bproc/patches/bproc-patch-2.4.17
\end{verbatim}

There may be other Linux patches in the patches directory with names
like \texttt{linux-\emph{x.y.z}-\emph{something}.patch}.  These
patches fix problems with Linux which affect BProc.  They are Linux
bugfixes, not Linux modifications for BProc.  These should be applied
in addition to the BProc kernel patch.

When configuring your kernel turn on \textbf{Beowulf Distributed
Process Space Support} under \textbf{General Setup}.  This will cause
the hooks for BProc to be built into the kernel.  The BProc kernel
code exists in kernel modules which will be built in the next step.

Build the new kernel and modules and install it on your system.
Modules built against another kernel will \textbf{not} work properly
with the new BProc kernel.  Several key structures are modified for
BProc and existing kernel modules will likely disfunction even if they
are still loadable.

Install the new kernel and reboot with it now.

\subsection*{Build and Install BProc}
Build the BProc source by typing \texttt{make} in the top level of the
BProc source tree.  The make files default to building against the
linux kernel tree in \texttt{/usr/src/linux}.  If your source tree is
in another location, add \texttt{LINUX=} to the \texttt{make} command
line.  For example:
\begin{verbatim}
make LINUX=/path/to/linux/source
\end{verbatim}
Next, do the install step as follows:
\begin{verbatim}
make LINUX=/path/to/linux/source install
\end{verbatim}

This will install all the user level programs, kernel modules,
libraries, and documentation on the system.

\subsection*{Configure BProc}


\subsubsection*{Run it all}
Install the kernel modules
\begin{verbatim}
modprobe bproc
bpmaster
\end{verbatim}


